---
title: "SISTRANGAS daily nominations"
subtitle: "qt11_visual"
format: html
editor: visual
---

# Data

```{r}
#| echo: false
#| results: hide
library(pacman)

# Use pacman to load and install the required packages
pacman::p_load(
  dplyr,
  flextable,
  ggplot2,
  htmltools,
  kableExtra,
  knitr,
  lubridate,
  magrittr,
  officer,
  purrr,
  RColorBrewer,
  readr, # For reading and writing data
  readxl,   # For reading Excel files
  scales,
  skimr,
  stringr,
  tidyverse,
  tidyr,
  webshot2,
  writexl
  )
```

```{r}
#| echo: false
#| results: hide

# 1 - Define a Custom Theme Function
theme_bloom <- function(base_size = 10, base_family = "sans") {
  theme_minimal(base_size = base_size, base_family = base_family) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 10, margin = margin(b = 10)),
      strip.text = element_text(face = "bold"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 5),

      # Bloomberg-style y-axis (right only)
      axis.title.y = element_blank(),
      axis.text.y.left = element_blank(),
      axis.ticks.y.left = element_blank(),
      axis.line.y.left = element_blank(),

      axis.text.y.right = element_text(),
      axis.ticks.y.right = element_line(),
      axis.line.y.right = element_line(),

      legend.position = "bottom",  # 👈 moved here
      legend.direction = "horizontal",
      legend.justification = "center",
      legend.box = "horizontal",
      legend.title = element_text(face = "bold", size = 8),  # 👈 smaller title
      legend.text = element_text(size = 7),                  # 👈 smaller labels
      legend.margin = margin(t = 10),
      panel.grid.major.x = element_blank(),
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_line(color = "grey85", linewidth = 0.2),  # 👈 lighter, thinner
      panel.grid.minor.y = element_line(color = "grey90", linewidth = 0.15),   # 👈 optional minor grid

      plot.caption = element_text(size = 7, hjust = 0, face = "italic", margin = margin(t = 10)),
      plot.margin = margin(15, 20, 15, 20)  # top, right, bottom, left
    )
}

# 2 - Define a Custom Color Scale Function
# Parametrized Version of scale_color_bloom()
# Default colors
# scale_color_bloom()
#
# Custom colors
# scale_color_bloom(values = c("grey30", "seagreen"))
# 
# Pass additional options 
# scale_color_bloom(values = c("black", "darkorange"),name = "Tipo de operación")
#
# "dodgerblue"
scale_color_bloom <- function(values = c("black", "turquoise"), ...) {
  scale_color_manual(values = values, ...)
}

# 3 - Define a Custom Date Scale Function
# Use default:
# scale_x_sistra_date()
#
# Custom breaks:
# scale_x_sistra_date(breaks = as.Date(c("2017-01-01", "2019-01-01", "2021-01-01")))
#
# Custom labels:
# scale_x_sistra_date(breaks = as.Date(c("2018-06-01", "2020-06-01")),labels = c("Mid-2018", "Mid-2020"))
#
scale_x_bloom <- function(breaks = NULL, labels = NULL, ...) {
  # Default breaks and labels if none provided
  if (is.null(breaks)) {
    breaks <- as.Date(c("2025-01-01", "2025-02-01", "2025-03-01", "2025-04-01"))
  }
  if (is.null(labels)) {
    labels <- format(breaks, "%Y-%m")
  }
  
  scale_x_date(breaks = breaks, labels = labels, ...)
}

# 4 reusable scale function
#
#   
scale_y_bloom <- function() {
  scale_y_continuous(position = "right")
}

# 5 make builder
#
make_bloom_plot <- function(data, x, y, color = NULL,
                            title = "Título del gráfico",
                            subtitle = "Subtítulo opcional",
                            caption = NULL,  # 👈 new argument
                            color_values = c("black", "turquoise"),
                            x_breaks = NULL, x_labels = NULL,
                            geom_type = c("point", "line")) {

  # Handle inputs
  geom_type <- match.arg(geom_type)
  color_quo <- rlang::enquo(color)
 
  # Start base ggplot
  p <- ggplot(data, aes(x = {{ x }}, y = {{ y }}, color = !!color_quo))

  # Add geom
  if (geom_type == "point") {
    p <- p + geom_point(size = 0.2, alpha = 0.5)
  } else if (geom_type == "line") {
    p <- p + geom_line(size = 0.4, alpha = 0.8)
  }

  # Add layers
  p <- p +
    labs(
      title = title,
      subtitle = subtitle,
       caption = caption,  # 👈 added here
      color = if (!rlang::quo_is_null(color_quo)) rlang::as_name(color_quo) else NULL
    ) +
    scale_color_bloom(values = color_values) +
    scale_x_bloom(breaks = x_breaks, labels = x_labels) +
    scale_y_bloom() +
    theme_bloom() +
    guides(x = guide_axis(n.dodge = 1))

  return(p)
}

```

```{r}
#| echo: false
#| results: hide

locations <- readRDS("qt_inputs/qt11_inputs/locations.rds")

str(locations)

loc_purp <- unique(locations$Loc.Purp)
```

```{r}
#| echo: false
#| results: hide

loc_rec <- locations %>%
  filter(Loc.Purp == "M2")

str(loc_rec)
```

## Locations info

In the next @tbl-Sistrangas-receipt-locations all the receipt locations connected to SISTRANGAS are enlisted.

```{r}
#| label: tbl-Sistrangas-receipt-locations
#| fig-width: 5
#| fig-asp: 0.7
#| echo: false
#| tbl-cap: SISTRANGAS locations

####
#### CHOICE 1
label_style <- "es"  # Choose: "es", "long", or "short"
####
####

# Source type translations
source_labels <- list(
  long = c(
    plant = "Processing plant",
    field = "Production field",
    border = "Border crossing",
    imp_pipe = "Pipeline interconnection",
    lng_terminal = "LNG terminal",
    virtual = "Virtual point",
    makeup = "Make-up gas",
    None = "Other"
  ),
  short = c(
    plant = "Plant",
    field = "Field",
    border = "Border",
    imp_pipe = "Interconn.",
    lng_terminal = "LNG",
    virtual = "Virtual",
    makeup = "Make-up",
    None = "Other"
  ),
  es = c(
    plant = "Planta de procesamiento",
    field = "Campo de producción",
    border = "Cruce fronterizo",
    imp_pipe = "Interconexión ducto",
    lng_terminal = "Terminal GNL",
    virtual = "Punto virtual",
    makeup = "Gas de reposición",
    None = "Otro"
  )
)

# Column names per language
column_labels <- list(
  long = c("Source type", "Node", "Name", "Design capacity (GJ/d)"),
  short = c("Type", "Node", "Name", "Cap. (GJ/d)"),
  es = c("Tipo de fuente", "Nodo", "Nombre", "Capacidad de diseño (GJ/d)")
)

# Caption per language
caption_text <- ifelse(
  label_style == "es",
  "Puntos de recepción en SISTRANGAS",
  "SISTRANGAS receipt locations"
)

# Transform data
loc_rec <- loc_rec %>%
  mutate(
    source_type = case_when(
      Loc %in% c("V025", "V905", "V907", "V908", "V9B6") ~ "plant",
      Loc %in% c("V028", "V036", "V042", "V045", "V051", "V056", "V059", "V077") ~ "field",
      Loc %in% c("V032", "V033", "V034", "V061", "V067") ~ "border",
      Loc %in% c("V030", "V037", "V062", "V074", "V076", "V078", "V918") ~ "imp_pipe",
      Loc %in% c("V055") ~ "lng_terminal",
      Loc %in% c("V071", "V072", "V073", "V075", "V079") ~ "virtual",
      Loc %in% c("MAKEUP 1", "MAKEUP 3", "MAKEUP 4", "MAKEUP 5", "MAKEUP 6", "MAKEUP 7", "MAKEUP 8") ~ "makeup",
      TRUE ~ "None"
    ),
    source_type = source_labels[[label_style]][source_type],
    Design.Capacity = comma(Design.Capacity, accuracy = 1)
  ) %>%
  arrange(source_type) %>%
  select(source_type, Loc, Loc.Name, Design.Capacity)

# Render stylized table
if (nrow(loc_rec) > 0) {
  kbl(
    loc_rec,
    align = "l",
    caption = caption_text,
    col.names = column_labels[[label_style]]
  ) %>%
    kable_styling(
      bootstrap_options = c("condensed"),
      full_width = TRUE,
      font_size = 12) %>%
    column_spec(1, bold = TRUE, color = "gray40", width = "10em") %>%
    column_spec(2, width = "6em") %>%
    column_spec(3, width = "12em") %>%
    column_spec(4, extra_css = "text-align: right;", width = "8em") %>%  # ✅ correct
    row_spec(0, bold = TRUE, color = "dodgerblue", extra_css = "border-bottom: 1px solid #DDD;") %>%
    row_spec(1:nrow(loc_rec), bold = FALSE, extra_css = "border-bottom: 1px solid #EEE;")
} else {
  cat("⚠️ No data available.")
}


# Convert your data frame to a flextable object
ft <- flextable(loc_rec)

# Create and export the Word document
doc <- read_docx() %>%
  body_add_par("My Table Title", style = "heading 2") %>%
  body_add_flextable(ft)

# Save to Word file
print(doc, target = "qt_outputs/qt11_outputs/table_output.docx")


```

### Scheduled and confirmed quantities

The chart displays daily nominations under the NAESB cycle (dots) and the actual scheduled quantities (grey line) for five key injection points. This information is critical for assessing operational imbalances within the gas pipeline network.

Scheduled quantities represent the flows that the pipeline operator has authorized for injection at each point. These injections must correspond either to:

-   nominated volumes at downstream delivery points, or

-   previously settled imbalance payments (e.g., over-deliveries or paybacks).

Failing this correspondence, the injection creates a new operational imbalance that must be managed through balancing mechanisms such as tolerance bands, make-up withdrawals, or imbalance charges, depending on the pipeline's balancing agreement.

```{r}
#| echo: false
#| results: hide

scheduling <- readRDS("qt_inputs/qt11_inputs/scheduling.rds")
nomination <- readRDS("qt_inputs/qt11_inputs/nomination.rds")

```

In the plot shown in @fig-Sistrangas-schedules-in-receipt-locations, each point represents the total scheduled quantity to be receipt in the next effective day.

```{r}
#| label: fig-Sistrangas-schedules-in-receipt-locations
#| fig-width: 8
#| fig-asp: 0.8
#| fig-cap: SISTRANGAS scheduled quantities
#| echo: false

# Caption per language
# Define caption dictionary
captions <- list(
  title01 = list(
    es = "Flujos de energía nominados por punto (TJ/día)",
    long = "Nominated Energy Flows by Point (TJ/day)",
    short = "Nom. Energy by point (TJ/d)"
  ),
  title02 = list(
    es = "Flujos de energía programados por punto (TJ/día)",
    long = "Scheduled Energy Flows by Point (TJ/day)",
    short = "Sched. Energy by point (TJ/d)"
  ),
  subtitle = list(
    es = "Recepción y entrega de gas",
    long = "Receipt and deliveries",
    short = "Rec/Del"
  ),
   name = list(
    es = "Ciclo",
    long = "Cycle",
    short = "Cycle"
  ),
  x_eje = list(
    es = "Día",
    long = "Day",
    short = "Day"
  )
)

# Select captions dynamically
capt_title01 <- captions$title01[[label_style]]
capt_title02 <- captions$title02[[label_style]]
capt_subt  <- captions$subtitle[[label_style]]
capt_name  <- captions$name[[label_style]]
capt_x  <- captions$x_eje[[label_style]]

p <-ggplot(data = nomination, aes(x = as.Date(Eff.Gas.Day.Time), y = Total.Scheduled.Quantity, color = Cycle.Desc)) +
  geom_point(size = 0.6, alpha = 0.9) +
  scale_color_manual(
    values = c("Timely" = "dodgerblue4", "Evening" = "dodgerblue", "Intraday 1" = "seagreen1", "Intraday 2" = "seagreen3", "Intraday 3" = "seagreen"),
    name = paste(capt_name)) +
    labs(title = paste(capt_title01),
         subtitle = paste(capt_subt),
       x = paste(capt_x)) +
  theme_bloom() +
  scale_x_bloom() +
  scale_y_bloom()

print(p)
```

In the plot shown in the next @fig-Sistrangas-confirmed-in-receipt-locations, it is possible to see the programmed quantities effective to each day. It matches with the scheduled values at the latest nomination cycle.

```{r}
#| label: fig-Sistrangas-confirmed-in-receipt-locations
#| fig-width: 8
#| fig-asp: 0.8
#| fig-cap: SISTRANGAS confirmed quantities
#| echo: false
p <-ggplot(data = scheduling, aes(x = Eff.Gas.Day, y = Total.Scheduled.Quantity)) +
  geom_point(color = "black", size = 0.5) +
  labs(title = paste(capt_title02),
         subtitle = paste(capt_subt),
       x = paste(capt_x)) +
  theme_bloom() +
  scale_x_bloom() +
  scale_y_bloom()

print(p)


```

# General Survey

## Location sample

1 plant : Cactus 1 border : Ramones 1 field : Mendoza 1 imp-pipe : Gaza (Gasoducto Zapotlanejo)

```{r}
#| echo: false
smp <- loc_rec %>%
  filter(Loc %in% c("V025", "V061", "V045","V076" )) %>%
  select(2)

# Filter for that nodo
nomin_smp <- nomination %>%
  inner_join(smp, by = "Loc")

# Get max cycle per date (for that nodo)
sched_smp <- scheduling %>%
  inner_join(smp, by = "Loc")

```

### Plot

@fig-Sistrangas-location

```{r}
#| label: fig-Sistrangas-location
#| fig-width: 8
#| fig-asp: 0.8
#| fig-cap: SISTRANGAS' Location sample
#| echo: false

captions <- list(
  title01 = list(
    es = "Energía programada por punto (TJ/día)",
    long = "Scheduled Energy Flows by Receipt Point (TJ/day)",
    short = "Sched. Energy by point (TJ/d)"
  ),
  subtitle = list(
    es = "Nodos clave de entrada: Cactus, Mendoza, Ramones, Gaza",
    long = "Key Entry Nodes: Cactus, Mendoza, Ramones, Gaza",
    short = "Nodos clave"
  ),
   name = list(
    es = "Ciclo",
    long = "Cycle",
    short = "Cycle"
  ),
  x_eje = list(
    es = "Día",
    long = "Day",
    short = "Day"
  )
)

# Select captions dynamically
capt_title01 <- captions$title01[[label_style]]
capt_subt  <- captions$subtitle[[label_style]]
capt_name  <- captions$name[[label_style]]
capt_x  <- captions$x_eje[[label_style]]

p <-ggplot(data = nomin_smp, aes(x = as.Date(Eff.Gas.Day.Time), y = Total.Scheduled.Quantity)) +
  geom_point(aes(color = factor(Cycle.Desc)), size = 0.6, alpha = 0.9) +
  geom_line(data = sched_smp, aes(x = Eff.Gas.Day, y = Total.Scheduled.Quantity), color = "gray60", size = 0.5) +
  scale_color_manual(
    values = c("Timely" = "dodgerblue4", "Evening" = "dodgerblue", "Intraday 1" = "seagreen1", "Intraday 2" = "seagreen3", "Intraday 3" = "seagreen"),
    name = paste(capt_name)) +
    labs(title = paste(capt_title01),
         subtitle = paste(capt_subt),
       x = paste(capt_x)) +
  theme_bloom() +
  scale_x_bloom() +
  scale_y_bloom() +
   facet_wrap(~Loc, ncol = 2, scales = "free_y")

print(p)

# Save the plot with high quality
ggsave("qt_outputs/qt11_outputs/images/sistra_keyplot.png", plot = p,
       width = 8, height = 4, dpi = 300, units = "in", bg = "white")

```

## Makeups by Tariff Zone

```{r}
#| echo: false

smp <- loc_rec %>%
  filter(Loc %in% c("MAKEUP 3", "MAKEUP 4", "MAKEUP 6","MAKEUP 8")) %>%
  select(2)


# Filter for that nodo
nomin_smp <- nomination %>%
  inner_join(smp, by = "Loc")

# Get max cycle per date (for that nodo)
sched_smp <- scheduling %>%
  inner_join(smp, by = "Loc")

```

### Plot

The next @fig-Sistrangas-make-ups

```{r}
#| label: fig-Sistrangas-make-ups
#| fig-width: 8
#| fig-asp: 0.8
#| fig-cap: Gas Make-up Volumes by Tariff Zone in SISTRANGAS
#| echo: false

captions <- list(
  title01 = list(
    es = "Seguimiento de gas de reposición por zonas tarifarias (TJ/día)",
    long = "Make-up Gas Tracking Across Tariff Zones (TJ/day)",
    short = "Make-ups by point (TJ/d)"
  ),
  subtitle = list(
    es = "Nominaciones y programaciones: Zonas 3, 4, 6 y 8",
    long = "Nominations and Schedules: Zones 3, 4, 6, 8",
    short = "Zones 3, 4, 6, 8"
  ),
   name = list(
    es = "Ciclo",
    long = "Cycle",
    short = "Cycle"
  ),
  x_eje = list(
    es = "Día",
    long = "Day",
    short = "Day"
  )
)

# Select captions dynamically
capt_title01 <- captions$title01[[label_style]]
capt_subt  <- captions$subtitle[[label_style]]
capt_name  <- captions$name[[label_style]]
capt_x  <- captions$x_eje[[label_style]]


p <-ggplot(data = nomin_smp, aes(x = as.Date(Eff.Gas.Day.Time), y = Total.Scheduled.Quantity)) +
  geom_point(aes(color = factor(Cycle.Desc)), size = 0.6, alpha = 0.9) +
  geom_line(data = sched_smp, aes(x = Eff.Gas.Day, y = Total.Scheduled.Quantity), color = "gray60", size = 0.5) +
  scale_color_manual(
    values = c("Timely" = "dodgerblue4", "Evening" = "dodgerblue", "Intraday 1" = "seagreen1", "Intraday 2" = "seagreen3", "Intraday 3" = "seagreen"),
    name = paste(capt_name)) +
    labs(title = paste(capt_title01),
         subtitle = paste(capt_subt),
       x = paste(capt_x)) +
  theme_bloom() +
  scale_x_bloom() +
  scale_y_bloom() +
   facet_wrap(~Loc, ncol = 2, scales = "free_y")

print(p)

# Save the plot with high quality
ggsave("qt_outputs/qt11_outputs/images/sistra_keyplot.png", plot = p,
       width = 8, height = 4, dpi = 300, units = "in", bg = "white")

```

```{r}
#| echo: false
# Get all unique source types
source_types <- unique(loc_rec$source_type)


captions <- list(
  title01 = list(
    es = "Energía programada (TJ/día)",
    long = "Scheduled Energy (TJ/day)",
    short = "Sched. Ene. (TJ/d)"
  ),
  subtitle = list(
    es = "Desglose por fuente",
    long = "Breakdown by Source",
    short = "Breakdown"
  ),
   name = list(
    es = "Ciclo",
    long = "Cycle",
    short = "Cycle"
  ),
  x_eje = list(
    es = "Día",
    long = "Day",
    short = "Day"
  )
)

# Select captions dynamically
capt_title01 <- captions$title01[[label_style]]
capt_subt  <- captions$subtitle[[label_style]]
capt_name  <- captions$name[[label_style]]
capt_x  <- captions$x_eje[[label_style]]



# Create a named list of plots
plots_by_type <- map(source_types, function(type) {

  # Filter locs for current type
  smp <- loc_rec %>%
    filter(source_type == type) %>%
    select(Loc)

  # Filter nomination and scheduling
  nomin_smp <- nomination %>%
    inner_join(smp, by = "Loc")

  sched_smp <- scheduling %>%
    inner_join(smp, by = "Loc")

########  

  # Generate the plot
  p <- ggplot(data = nomin_smp, aes(x = as.Date(Eff.Gas.Day.Time), y = Total.Scheduled.Quantity)) +
    geom_point(aes(color = factor(Cycle.Desc)), size = 0.6, alpha = 0.9) +
    geom_line(data = sched_smp, aes(x = Eff.Gas.Day, y = Total.Scheduled.Quantity), color = "gray60", alpha = 0.6, size = 0.5) +
    scale_color_manual(
    values = c("Timely" = "dodgerblue4", "Evening" = "dodgerblue", "Intraday 1" = "seagreen1", "Intraday 2" = "seagreen3", "Intraday 3" = "seagreen"),
    name = paste(capt_name)) +
    labs(title = paste(capt_title01),
         subtitle = paste(capt_subt, ": ", type),
       x = paste(capt_x)) +
  theme_bloom() +
  scale_x_bloom() +
  scale_y_bloom() +
  facet_wrap(~Loc, ncol = 3, scales = "free_y")

  return(p)
})

# Name the list for easier access
names(plots_by_type) <- source_types


```

```{r}
#| echo: false
# label_style should be one of: "es", "long", "short"
tr <- function(es, long, short) {
  if (label_style == "es") {
    es
  } else if (label_style == "long") {
    long
  } else {
    short
  }
}

# Correct argument order: (es, long, short)
filter_01 <- tr(
  "Planta de procesamiento",  # Spanish
  "Processing plant",         # English long
  "Plant"                     # English short
)

filter_02 <- tr(
  "Campo de producción",  # Spanish
  "Production field",         # English long
  "Field"                     # English short
)

filter_03 <- tr(
  "Cruce fronterizo",  # Spanish
  "Border crossing",         # English long
  "Border"                     # English short
)

filter_04 <- tr(
  "Interconexión ducto",  # Spanish
  "Pipeline interconnection",         # English long
  "Interconn."                     # English short
)

filter_05 <- tr(
  "Terminal GNL",  # Spanish
  "LNG terminal",         # English long
  "LNG"                     # English short
)

 filter_06 <- tr(
  "Punto virtual",  # Spanish
  "Virtual point",         # English long
  "Virtual"                     # English short
)
 
 filter_07 <- tr(
  "Gas de reposición",  # Spanish
  "Make-up gas",         # English long
  "Make-up"                     # English short
)
```

```{r}
#| echo: false
# Access the plot
plots_by_type[[filter_01]]

```

```{r}
#| echo: false
# Access the plot
plots_by_type[[filter_02]]

```

```{r}
#| echo: false
# Access the plot
plots_by_type[[filter_03]]

```

```{r}
#| echo: false
# Access the plot
plots_by_type[[filter_04]]
```

```{r}
#| echo: false
# Access the plot
plots_by_type[[filter_05]]
```

```{r}
#| echo: false
#| # Access the plot
plots_by_type[[filter_06]]
```

# Today's report

```{r}
#| echo: false
# Get all unique source types
sched.last.day <- scheduling %>%
  filter(Eff.Gas.Day == max(unique(scheduling$Eff.Gas.Day)))

# Create a named list of plots
sched_by_type <- map(source_types, function(type) {

  # Filter locs for current type
  smp <- loc_rec %>%
    filter(source_type == type) %>%
    select(Loc, Loc.Name)

  sched_smp <- sched.last.day %>%
    inner_join(smp, by = "Loc")

  # Generate the plot
  p <- ggplot(data = sched_smp, aes(x = Loc.Name, y = Total.Scheduled.Quantity)) +
       geom_bar(stat = "identity", fill = "dodgerblue") +
    labs(title = paste("Scheduled Energy on", max(unique(scheduling$Eff.Gas.Day)) ,"(TJ/day)"),
         subtitle = paste("Breakdown by ", type, "Location"),
       x = "Location") +
  theme_bloom() +
  scale_y_bloom()
  
  return(p)
})

# Name the list for easier access
names(sched_by_type) <- source_types

```

```{r}
#| echo: false
sched_by_type[[filter_01]]

```

```{r}
#| echo: false
sched_by_type[[filter_02]]

```

```{r}
#| echo: false
sched_by_type[[filter_03]]

```

```{r}
#| echo: false
sched_by_type[[filter_04]]

```

# Filter Makeups and virtual

```{r}
#| echo: false
loc_effe <- loc_rec %>%
  filter(!source_type %in% c(filter_06, filter_07))

saveRDS(loc_effe, file = "qt_outputs/qt11_outputs/loc_effe.rds")
```
