---
title: "qt12_02_maps"
format: html
---

# Check previous Maps

```{r}
#| echo: false
#| results: hide
library(pacman)

# Use pacman to load and install the required packages
pacman::p_load(
  arrow,
  dplyr,
  flextable,
  ggplot2,
  ggspatial,
  htmltools,
  htmlwidgets,
  kableExtra,
  knitr,
  leaflet,
  lubridate,
  magrittr,
  mapview,
  officer,
  purrr,
  RColorBrewer,
  readr, # For reading and writing data
  readxl,   # For reading Excel files
  rnaturalearth,
  rnaturalearthdata,
  scales,
  skimr,
  sf,
  stringr,
  tidyverse,
  tidyr,
  viridis,
  webshot2,
  writexl
  )
```

```{r}
#| echo: false
#| results: hide

base_colors <- c("black","bisque4","steelblue3", "slategray2","seagreen3")
base_tono <- c("steelblue1", "bisque","seagreen3","grey23")

```

```{r}
#| echo: false

# points
injection_map <- readRDS("qt_outputs/qt12_outputs/injection_map.rds")

injection_map

```

# New map

## Natural gas injection points

This map displays the geographic distribution of specific natural gas injection nodes across Mexico's national pipeline system. Each marker corresponds to a unique point of receipt, identified by its node code (e.g., V025, V030), and colored individually for distinction. The legend in the bottom-left matches each color to its respective node code. This layout supports operational mapping and node-level analysis, offering a clear spatial reference for infrastructure planning, balancing strategies, or supply source monitoring across the SISTRANGAS network.

```{r}
#| echo: false
#| results: hide

# points
injection_map01 <- readRDS("qt_outputs/qt12_outputs/injection_map01.rds")

injection_map01

```

## Natural gas injection point along the network

```{r}
#| echo: false
#| results: hide
#| message: false
# Read the KML file directly from the path
sistra <- st_read("qt_inputs/qt12_inputs/tra_kml/SISTRANGAS.kml") %>%
  st_zm(drop = TRUE, what = "ZM")

```

```{r}
#| echo: false
# Create and display the map with leaflet
# 0 = use leaflet_map01
# 1 = use leaflet_map
use_main_map <- 0  # 👈 change this to 0 or 1 as needed

# Select base map based on the binary variable
selected_map <- if (use_main_map == 1) injection_map else injection_map01

# Now apply your conditional logic
if (exists("sistra")) {
  leaflet_map02 <- selected_map %>%
    addPolylines(data = sistra, color = "dodgerblue", weight = 1)
}

# Display result
leaflet_map02

```

```{r}
#| echo: false
#| results: hide
saveRDS(leaflet_map02, "qt_outputs/qt12_outputs/injection_map02.rds")

```

```{r}
#| echo: false

### Cambio a Verdana

leaflet_map02 <- leaflet_map02 %>%
  htmlwidgets::onRender("
    function(el, x) {
      var style = document.createElement('style');
      style.innerHTML = `
        .leaflet-popup-content, .leaflet-control {
          font-family: Verdana, sans-serif;
          font-size: 12px;
        }
      `;
      document.head.appendChild(style);
    }
  ")

leaflet_map02
```

```{r}
#| echo: false
#| results: hide

saveWidget(leaflet_map02, "qt_outputs/qt12_outputs/injection_map02.html", selfcontained = TRUE)
# If not installed:
# install.packages("webshot2")
mapview::mapshot(
  x = leaflet_map02,
  file = "qt_outputs/qt12_outputs/injection_map02.png",
  selfcontained = TRUE,
  vwidth = 1200,       # Width in pixels
  vheight = 900       # Height in pixels
)
```
